---
title: "DataCamp-Intro to R"
author: "Katherine"
date: "July 16, 2018"
output: html_document
---
# Inro to R
```{r}
##Arithmatic with R
4+6
4-6
4*6
4/2
2^3
###Modulo
6%%4  
-9%%6
```

```{r}
##Variable Assignment
x <- 42
x
my_apples <- 5
my_oranges <- 6
my_fruit <- my_apples+my_oranges
my_fruit
```

```{r}
##Basic Data Types in R
my_numeric <- 42
my_character <- "universe"
my_logical <- FALSE
class(my_numeric)
class(my_character)
class(my_logical)
```

##Vectors
```{r}
vegas <- "GO!"
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")
boolean_vector <- c(TRUE, FALSE, TRUE)

poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)

### name vector 1
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(roulette_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
### name vector 2
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

### calculating total winnings
total_vector <- poker_vector+roulette_vector
total_vector
total_poker <- sum(poker_vector)
total_roulette <- sum(roulette_vector)
total_week <- sum(total_poker+total_roulette)
total_week

### comparing total winnings
total_poker > total_roulette

### vector selection
poker_wednesday <- poker_vector[3] 
poker_midweek <- poker_vector [c(2,3,4)]
roulette_selection_vector <- roulette_vector[2:5]
poker_start <- poker_vector[c("Monday", "Tuesday", "Wednesday")]
mean(poker_start)

### selection by comparison
selection_vector <- poker_vector>0
selection_vector
poker_winning_days <- poker_vector[selection_vector]
poker_winning_days
selection_vector2 <- roulette_vector>0
selection_vector2
roulette_winning_days <- roulette_vector[selection_vector2]
roulette_winning_days
```

##Matrix
```{r}
matrix(1:9, byrow=TRUE, nrow=3)
matrix

### box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

box_office <- c(new_hope, empire_strikes, return_jedi)
star_wars_matrix <- matrix(box_office, byrow=TRUE, nrow=3)

### naming a matrix
###rownames(my_matrix) <- row_names_vector
###colnames(my_matrix) <- col_names_vector
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")
colnames(star_wars_matrix) <- region
rownames(star_wars_matrix) <- titles
star_wars_matrix

### calculating the worldwide box office- rowSums, cbind
worldwide_vector <- rowSums(star_wars_matrix)
worldwide_vector
all_wars_matrix <- cbind(star_wars_matrix, worldwide_vector)
all_wars_matrix

### constr. star_wars_matrix2
box_office2 <- c(474.5, 552.5, 310.7, 338.7, 380.3, 468.5)
star_wars_matrix2 <- matrix(box_office2, nrow=3, byrow=TRUE)
star_wars_matrix2
titles2 <- c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")
colnames(star_wars_matrix2) <- region
rownames(star_wars_matrix2) <- titles2
star_wars_matrix2

### combine star_wars_matrix & star_wars_matrix2- rbind
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
all_wars_matrix

### colSums
total_revenue_vector <- colSums(all_wars_matrix)
total_revenue_vector

### selection of matrix elements
non_us_all <- all_wars_matrix[,2]
mean(non_us_all)
non_us_some <- all_wars_matrix[1:2,2]
mean(non_us_some)

### arithmetic with matrices
visitors <- all_wars_matrix/5
visitors

### ticket prices matrix
tickect_prices <- c(5.0, 5.0, 6.0, 6.0, 7.0,7.0, 4.0,4.0, 4.5, 4.5, 4.9, 4.9)
ticket_prices_matrix <- matrix(tickect_prices, byrow=TRUE, nrow=6)
ticket_prices_matrix
titles3 <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi", "The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")
colnames(ticket_prices_matrix) <- region
rownames(ticket_prices_matrix) <- titles3
ticket_prices_matrix

###vistors with diff. ticket prices
visitors= all_wars_matrix/ ticket_prices_matrix
visitors
us_visitors <- visitors[,1]
mean(us_visitors)
```

##Factor:The term factor refers to a statistical data type used to store categorical variables.
```{r}
sex_vector <- c("Male", "Female", "Female", "Male", "Male")
factor_sex_vector <- factor(sex_vector)
factor_sex_vector

```
###There are two types of categorical variables: a nominal categorical variable and an ordinal categorical variable.
####A nominal variable is a categorical variable without an implied order. In contrast, ordinal variables do have a natural ordering. 
```{r}
### Animals - Categorical variables
animals_vector <- c("Elephant", "Giraffe", "Donkey", "Horse")
factor_animals_vector <- factor(animals_vector)
factor_animals_vector

### Temperature - Ordinal variables
temperature_vector <- c("High", "Low", "High","Low", "Medium")
factor_temperature_vector <- factor(temperature_vector, order = TRUE, levels = c("Low", "Medium", "High"))
factor_temperature_vector

### Factor Levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")
factor_survey_vector
summary(survey_vector)
summary(factor_survey_vector)
### comparing with the sexes - meaningless
male <- factor_survey_vector[1]
female <- factor_survey_vector[2]
male > female

###Ordinal variables
speed_vector <- c("medium","slow","slow","medium","fast")
factor_speed_vector <- factor(speed_vector, order = TRUE, levels= c("slow", "medium","fast"))
factor_speed_vector
summary(factor_speed_vector)
da2 <- factor_speed_vector [2]
da5 <- factor_speed_vector [5]
da2 > da5
```

##Data Frame
```{r}
mtcars
head(mtcars)
tail(mtcars)
str(mtcars)


name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)
planets_df <- data.frame(name, type, diameter, rotation, rings)
planets_df
str(planets_df)
### selection of df
#### Mercury diameter & Mars all 
planets_df[1,3] 
planets_df[4,]
#### Select first 5 values of diameter column
planets_df[1:5, "diameter"]
rings_vector <- planets_df$rings   ####planets_df[,"rings"]
rings_vector
planets_df[rings_vector,]
subset(planets_df, subset=diameter<1)

#### Sorting
a <- c(99, 9, 900)
order(a)
a[order(a)]

positions <- order(planets_df$diameter)
positions
planets_df[positions,]
```

##Lists
```{r}
my_vector <- 1:10 
my_matrix <- matrix(1:9, ncol=3)
my_df <- mtcars[1:10,]
my_list <- list(my_vector, my_matrix, my_df)
my_list <- list(vec=my_vector, mat=my_matrix, df=my_df)
my_list

### Lists selection
####3rd components 4th element
my_list[[3]][4]
my_list$vec
my_list[["mat"]]
my_list_full <- c(my_list, yeaar=2018)
my_list_full
str(my_list_full)
```
